---
title: "GSS challange"
author: "Precious Calderon"
date: "06-21-2022"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2018.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with The Most data (fewer variables gives more data points)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
Based on the graph showing the counts of people in favor and opposed to capital punishment, there is a significantly larger number of white people who are in favor of capital punishment compared to other racial groups. Meanwhile, about half of the number of white people who agreed to capital punishment oppose it. African Americans, on the other hand, generally oppose capital punishment while other racial groups appear to be in favor of it but not to the same extent as white people.  

The second graph, meanwhile, shows the percentage of people in each racial groups that are in favor and are opposed to capital punishment. It shows that more than 60% of white people agree to capital punishment while less than 40% are opposed to it. More than half of black people are opposed to capital punishment, and more than half of other racial groups are in favor of it. 
  
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->




## Now with More data

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```
  
The first table compares the percentages of people who are in favor of and are opposed to capital punishment per racial group. It specifies that 67.98% of white people are in favor of capital punishment while 32.02% are opposed to it. Meanwhile, 46.29% of black people are in favor of capitol punishment while 53.71% are opposed to it. For other racial groups, however, 55.87% are in favor while 44.13% are opposed. Therefore, the first table shows that white and other racial groups are generally in favor of capital punishment while black people are opposed to it.
  
The second table, on the other hand, shows the percentages of racial groups that make up the population of people who are in favor of and are opposed to capitol punishment, respectively. Because there are more white people compared to other racial groups, they make up most of the population who are in favor and are opposed to capitol punishment, with 78.34% and 63.24%, respectively. And because there are less number of black people and others in the sample, they make up less of the population who are in favor and are opposed to capitol punishment.  


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`


## More data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```

```{r}
fisherXY(mydata$cappun,mydata$race)
```
 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
